---
title: "week3-E3"
author: "AureliaMo"
date: "2023-05-05"
output: html_document
---

## Exercise 3

libraries
```{r}
library(readr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(lubridate)

library(sf)
library(tmap)

library(SimilarityMeasures)
```

## Preparation

import data and transform it to the coordinate system. 

```{r}
mydata <- read_delim("aurelia_dataset.csv", delim = ";")

# create datetime column
mydata <- mydata |> 
  mutate(Time = as.character(Time),
         datetime = paste(Date,Time),
         datetime = as.POSIXct(datetime, format = "%d.%m.%Y %X", tz="UTC"),
         datetime = with_tz(datetime, tzone = "Europe/Berlin")
         ) |> 
  select(datetime, Latitude, Longitude)

# convert data to an sf object
mydata <- st_as_sf(mydata, coords = c("Longitude", "Latitude"), crs = 4326, remove = FALSE) # EPSG code for WGS84 = 4326

# convert data to coordinate system CH1903+ LV95
mydata <- st_transform(mydata, crs = 2056) 

head(mydata)
```

store the coordinates in separate columns, explore Data and filter the day i cycled to Waedenswil

```{r}
# store the coordinates in separate columns
mycoord <- st_coordinates(mydata)

mydata <- cbind(mydata, mycoord)

# explore Data
tmap_mode("view")
tm_shape(mydata)+
  tm_dots()

# choose a single day for the next steps 

oneday <- mydata |> 
  filter(as.Date(datetime) == "2023-04-06")

head(oneday)
tm_shape(oneday)+
  tm_dots()
```


## Task 1 Segmentation

Step a): Specify a temporal window 
I was riding my bike and the sampling interval was 5 seconds. If i didn't move for more than 20 secs, i probably waited at a red light, or for my friends to join me, or to check where i hat do go. So i set my temporal window to 20 secs and therefore to four fixes

 pos[n-2] to pos[n]
 pos[n-1] to pos[n]
 pos[n] to pos[n+1]
 pos[n] to pos[n+2]



Step b): Measure the distance from every point to every other point within this temporal window. Calculate the mean distance between the points. 

```{r}

oneday <- oneday |> 
  mutate(
    n_plus1 = sqrt((lead(X)-X)^2 + (lead(Y)-Y)^2),
    n_plus2 = sqrt((lead(X,2)-X)^2 + (lead(Y,2)-Y)^2), # new offset of 2 in function lead()
    n_minus1 = sqrt((lag(X)-X)^2 + (lag(Y)-Y)^2),
    n_minus2 = sqrt((lag(X,2)-X)^2 + (lag(Y,2)-Y)^2)
  )

oneday <- oneday |>
  rowwise() |>  # so the mean per row is computed
  mutate(
    stepmean = mean(c(n_plus1, n_plus2, n_minus1, n_minus2))
  ) |> 
  ungroup() # otherwise each row would represent a group, we don't want that

```

## Task 2: Specify and apply threshold d

exploring the euclidean distance values. how is the distribution? 
```{r}

ggplot(oneday, aes(stepmean))+
  geom_histogram(binwidth = 5)+
  geom_vline(xintercept = mean(oneday$stepmean, na.rm=TRUE))

# the distribution shows two peaks. one at lower mean steplength (5 m) and one around mean steplengths of 50m. Segmenting by the mean of the values seems reasonable. 

```

We use the mean of all stepmean values to separate moves from static points. Store the new information (boolean to differentiate between stops (TRUE) and moves (FALSE)) in a new column named static.

````{r}
oneday <- oneday |> 
  mutate(static = stepmean < mean(stepmean, na.rm = TRUE)) 

```


## Task 3: Visualize segmented trajectories

```{r}
ggplot(oneday, aes(X, Y))+
  geom_path()+
  geom_point(aes(color = static))+
  coord_fixed()
```

I walked the last bit from Waedenswil bahnhof to the campus, and there were some steep parts on the way to Oerlikon, where i moved slower. Also, I stopped sometimes to check the directions and wait for my friends. The visualisation represents my stops adequately i think.


## Task 4 Segment-based analysis

assign unique IDs to the segments

```{r}
# function for assigning unique ID to the subtrajectories

rle_id <- function(vec) {
    x <- rle(vec)$lengths
    as.factor(rep(seq_along(x), times = x))
}

oneday <- oneday |> 
  mutate(ID=rle_id(static))

head(oneday)
```

Visualize the moving segments by colourizing them by segment_ID.
```{r}

oneday <- oneday |> 
  filter(!static) # select rows where the static value is not TRUE to select the moving segments

head(oneday)

ggplot(oneday, aes(X, Y))+
  geom_point()+
  geom_path(aes(color=ID))+
  coord_fixed()

# gosh, i have many different segments!
```

Use segment_ID as a grouping variable to determine the segments duration and remove short segments (e.g. segments with a duration < 5 Minutes)

```{r}

# oneday$datetime[1]
# oneday$datetime[length(oneday$datetime)]
# 
#   
# as.numeric(difftime(oneday$datetime[1], oneday$datetime[length(oneday$datetime)]))
# 
# oneday |> 
#   group_by(ID) |> 
#   as.numeric(difftime(datetime[length(datetime)], datetime[1]))
# 
# head(oneday)
# haven't found out why it doesn't work yet

```



## Task 5: Similarity measures

import and explore
```{r}

pedestrian <- read_csv("pedestrian.csv")
head(pedestrian)

ggplot(pedestrian, aes(E, N))+
  geom_point(aes(color = TrajID))+
  geom_path(aes(color = TrajID))+
  facet_wrap(~ TrajID, nrow = 2)+
  theme_minimal()
```


## Task 6: Calculate similarity

install package, familiarize
```{r}
# install.packages("SimilarityMeasures")

  help(package = "SimilarityMeasures")

```

Now compare trajectory 1 to trajectories 2-6 using different similarity measures from the package. Your options are. DTW, EditDist, Frechet and LCSS.
Before visualizing your results think about the following: Which two trajectories to you percieve to be most similar, which are most dissimilar? hmm... Trajectory 3 and 6 look very similar to me. Trajectory 4 looks very dissimilar to all the others.

Now visualize the results from the computed similarity measures.
First, i need to create one matrix for every trajectory.

```{r}
# is the datetime column relevant? according to the input examples of the functions description i think not. 

pedestrian <- pedestrian |> 
  select(-DatetimeUTC)


traj1 <- pedestrian |> 
  filter(TrajID==1) |> 
  select(-TrajID) |> 
  as.matrix()
  
traj2 <- pedestrian |> 
  filter(TrajID==2) |> 
  select(-TrajID) |> 
  as.matrix()

traj3 <- pedestrian |> 
  filter(TrajID==3) |> 
  select(-TrajID) |> 
  as.matrix()

traj4 <- pedestrian |> 
  filter(TrajID==4) |> 
  select(-TrajID) |> 
  as.matrix()

traj5 <- pedestrian |> 
  filter(TrajID==5) |> 
  select(-TrajID) |> 
  as.matrix()

traj6 <- pedestrian |> 
  filter(TrajID==6) |> 
  select(-TrajID) |> 
  as.matrix()
```

compute similarity measures DTW, EditDist, Frechet and LCSS.
compare trajectory 1 to trajectories 2 to 6.

```{r}
DTW_comparison <- c(DTW(traj1, traj2),
DTW(traj1, traj3),
DTW(traj1, traj4),
DTW(traj1, traj5),
DTW(traj1, traj6))

EditDistance_comparison <- c(EditDist(traj1, traj2, pointDistance = 20),
EditDist(traj1, traj3),
EditDist(traj1, traj4),
EditDist(traj1, traj5),
EditDist(traj1, traj6))

Frechet_comparison <- c(Frechet(traj1, traj2, testLeash=-1),
Frechet(traj1, traj3),
Frechet(traj1, traj4),
Frechet(traj1, traj5),
Frechet(traj1, traj6))

LCSS_comparison <- c(LCSS(traj1, traj2, pointSpacing=3, pointDistance=20, 
     errorMarg=2, returnTrans=FALSE),
LCSS(traj1, traj3, pointSpacing=3, pointDistance=20, 
     errorMarg=2, returnTrans=FALSE),
LCSS(traj1, traj4, pointSpacing=3, pointDistance=20, 
     errorMarg=2, returnTrans=FALSE),
LCSS(traj1, traj5, pointSpacing=3, pointDistance=20, 
     errorMarg=2, returnTrans=FALSE),
LCSS(traj1, traj6, pointSpacing=3, pointDistance=20, 
     errorMarg=2, returnTrans=FALSE))

comparison <- data.frame(DTW_comparison, EditDistance_comparison, Frechet_comparison, LCSS_comparison)
comparison$traj <- c(1,2,3,4,5)

head(comparison)



ggplot(comparison, aes(traj, DTW_comparison, fill=traj))+
  geom_bar(stat = "identity")

ggplot(comparison, aes(traj, EditDistance_comparison, fill=traj))+
  geom_bar(stat = "identity")

ggplot(comparison, aes(traj, Frechet_comparison, fill=traj))+
  geom_bar(stat = "identity")

ggplot(comparison, aes(traj, LCSS_comparison, fill=traj))+
  geom_bar(stat = "identity")



```

